### 2021-02-23 architecture.tex
Divide Sorting into subsubsections

We extend the notion of cost from symbols to precedences
by summing up the symbol costs
weighted by their positions in the given precedence $\pi$:
$$
C(\pi) = Z_n \sum_{i=1}^n i \cdot c(\pi(i))
% We call the normalization factor Z_n to follow convention from Foundations of Machine Learning, p. 122.
$$
$Z_n = \frac{2}{n(n+1)}$ is a normalization factor
that ensures the commensurability of precedence costs across signature sizes.
More precisely, normalizing by $Z_n$ makes the expected value of precedence cost
independent of the signature size $n$:
\begin{align*}
\mathbb{E}_\pi [C(\pi)]
&= \mathbb{E}_\pi \SquareBracket{Z_n \sum_{i=1}^n i \cdot c(\pi(i))} \\
&= Z_n \sum_{i=1}^n i \cdot \mathbb{E}_{\pi, i} [c(\pi(i))] \\
&= Z_n \Parentheses{\sum_{i=1}^n i} \mathbb{E}_i [c(i)] \\
&= \frac{2}{n(n+1)} \frac{n(n+1)}{2} \mathbb{E}_i [c(i)] \\
&= \mathbb{E}_i [c(i)]
\end{align*}
\todo{FB: Consider making the background problem explicit by using $C(\pi|P)$ or $C(\pi,P)$.}
\todo[inline]{FB: Consider accessing elements $\pi$ as vector elements $\pi_i$ instead of $\pi(i)$. Keep in mind that $\inv{\pi}$ must be updated as well. The motivation for using $\pi(i)$ is that $\inv{\pi}(i)$ looks clearer than $\inv{\pi}_i$. How about using $(\inv{\pi})_i$? We may also omit the inversion completely.}

\begin{lemma}
Precedence cost $C$ is minimized by the precedence that sorts the symbols by their costs in non-increasing order:
$$
\argmin_\pi C(\pi) \in \argsort^- (c(1), \ldots, c(n))
$$
where $\argsort^-(x)$ is the set of all permutations $\rho$ that sort vector $x$ in non-increasing order ($x_{\rho(1)} \geq x_{\rho(2)} \geq \ldots \geq x_{\rho(n)}$).
\end{lemma}

>>>>

\subsubsection{Precedence cost}
We extend the notion of cost from symbols to precedences
by summing up the symbol costs
weighted by their positions in the given precedence $\pi$:
$$
C(\pi) = Z_n \sum_{i=1}^n i \cdot c(\pi(i))
% We call the normalization factor Z_n to follow convention from Foundations of Machine Learning, p. 122.
$$
$Z_n = \frac{2}{n(n+1)}$ is a normalization factor
that ensures the commensurability of precedence costs across signature sizes.
More precisely, normalizing by $Z_n$ makes the expected value of precedence cost
independent of the signature size $n$:
\begin{align*}
\mathbb{E}_\pi [C(\pi)]
&= \mathbb{E}_\pi \SquareBracket{Z_n \sum_{i=1}^n i \cdot c(\pi(i))} \\
&= Z_n \sum_{i=1}^n i \cdot \mathbb{E}_{\pi, i} [c(\pi(i))] \\
&= Z_n \Parentheses{\sum_{i=1}^n i} \mathbb{E}_i [c(i)] \\
&= \frac{2}{n(n+1)} \frac{n(n+1)}{2} \mathbb{E}_i [c(i)] \\
&= \mathbb{E}_i [c(i)]
\end{align*}
\todo{FB: Consider making the background problem explicit by using $C(\pi|P)$ or $C(\pi,P)$.}
\todo[inline]{FB: Consider accessing elements $\pi$ as vector elements $\pi_i$ instead of $\pi(i)$. Keep in mind that $\inv{\pi}$ must be updated as well. The motivation for using $\pi(i)$ is that $\inv{\pi}(i)$ looks clearer than $\inv{\pi}_i$. How about using $(\inv{\pi})_i$? We may also omit the inversion completely.}

\subsubsection{Cost minimization by sorting}
\begin{lemma}
Precedence cost $C$ is minimized by the precedence that sorts the symbols by their costs in non-increasing order:
$$
\argmin_\pi C(\pi) \in \argsort^- (c(1), \ldots, c(n))
$$
where $\argsort^-(x)$ is the set of all permutations $\rho$ that sort vector $x$ in non-increasing order ($x_{\rho(1)} \geq x_{\rho(2)} \geq \ldots \geq x_{\rho(n)}$).
\end{lemma}
